---
title: "Iteration"
author: "Kaleb J. Frierson"
date: "2024-10-29"
output: html_document
---

# Notes

for loops:

Loops are the easiest place to start with iteration. Loops have a consistent output object. The basic structure: 

input = list(...)
output = list(...)

for (i in 1:n) {
output [[i]] = f(input[[i]])

}

Mapping across input list, running function, saving output. 

Loop functions make your life easier. You can use them in the context of a tidyverse like approach. Makes your code easier to read. More focus on the things you are trying to do/change. Map function is in the purrr package and is a loop function that helps clean up loops. Don't use base R loop functions, they're terrible per Jeff. 

output = map(input, f)

This produces the same result as a for loop but is better because: [answer here]

Map variants: 

By default, map makes one input and will return a list. If you know what kind of output your function will produce you can change default settings. 

Process: 

Write a single example for fixed inputs. 
Abstract example to a function
Embed in a loop 
Re-write useing a map statement 

(IE build the car model)

Lists: they exist, now we will worry about it. 

R uses lists to keep track of anything you want to use. They're a collection of things. One list can have lots of different things in them. 

Data frames - special kind of list. 

Lists can contain lists. If you start with a df and have a new list column and want to put it into a df, you can put it in there. This is very helpful. Inside of your data frame you can have input list as a column. Keeping everything in one data frame with list columns means there are fewer things to worry about. 

# Time to Code

```{r, message=FALSE}
library(tidyverse)
library(rvest)
library(knitr)
```


## Here's some lists 

```{r}
l = list(
  vec_numeric = 1:4, 
  unif_sample = runif(100), 
  mat = matrix(1:8, nrow = 2, ncol = 4, byrow = TRUE), 
  summary = summary(rnorm(1000))
)

```

List more useful: 

```{r}
list_norm = 
  list(
    a = rnorm(20, 0, 5),
    b = rnorm(20, 4, 5), 
    c = rnorm(20, 0, 10), 
    d = rnorm(20, 4, 10)
  )

list_norm$b
```
lets reuse the function we wrote last time: 
```{r}
mean_and_sd = function(x) {
  
  mean_x = mean(x)
  sd_x = sd(x)
  
  out_df = 
    tibble(
      mean=mean_x, 
      sd = sd_x
    )
  
  return(out_df) 
}


```


Lets use the function to take the mean and sd of all samples: 

```{r}
mean_and_sd(list_norm$a)
mean_and_sd(list_norm$b)
mean_and_sd(list_norm$c)
mean_and_sd(list_norm$d) 
```

If I had 400 of these, it would not be easy to make sure this still worked using this way. That is what loops are for!

## Use a for loop

create output list and run a for loop: 

```{r}
output = vector("list", length = 4) 

for (i in 1:4) { 
  
  output[[i]] = mean_and_sd(list_norm[[i]])
  
  }

output 

```

## Do the same thing with map 

```{r}
output = map(list_norm, mean_and_sd) 

```

^ this line of code is the same thing that we did above. I think map is more intuitive but there are a lot of ways to do this. Map seems fantastic. 

Lets do a couple of other things: 

```{r}
output = map(list_norm, median)
output_iqr = map(list_norm, IQR)

output_iqr2 = map_dbl(list_norm, IQR)

output_rows = 
  map(list_norm, mean_and_sd) |> 
  bind_rows()
```

## List Columns!!

Lets get tidy : ) 

```{r}
listcol_df = 
  tibble(
    name = c("a", "b", "c", "d"), 
    samp = list_norm
  )

listcol_df |> 
  filter(name %in% c("a", "b"))

listcol_df |>  
  select(-samp)

```

This whole df is still a list! You can access list variable like we have done, see below. 
```{r}
listcol_df[["samp"]][["a"]]
```

compute mean and sd 
```{r}
mean_and_sd(listcol_df[["samp"]][["a"]])
```
Since its still just a list, I can do the same things I have done before. We can map over a list and compute a given function. 

```{r}
output_list = map(listcol_df[["samp"]], mean_and_sd)
output_list
```

Refresher - what is going on? 

The goal of map is to iterate across inputs, apply a function, and spit out results. We tell it what to input, what function to apply, and it automatically produces the output. Is it creating a new column? Not yet, you have to tell it to. 


Add list column: 
```{r}
listcol_df = 
  listcol_df|>
  mutate(output = map(samp, mean_and_sd), 
         iqr = map_dbl(samp, IQR)) 

listcol_df = 
  listcol_df|>
  mutate(output = map_dfr(samp, mean_and_sd), 
         iqr = map_dbl(samp, IQR)) 

listcol_df = 
  listcol_df|>
  mutate(output = map(samp, mean_and_sd), 
         iqr = map_dbl(samp, IQR)) |> 
  select(-samp) |> 
  unnest(output)

```

### NSDUH 

Want to have a df that includes the drug, extracts tables, and puts everything in one dataset. 

version of function from before: 
```{r}
nsduh_table_format = function(html, table_num) {
  
  out_table = 
    html |> 
    html_table() |> 
    nth(table_num) |> 
    slice(-1) |> 
    select(-contains("P Value")) 
  
  return(out_table)
}

```

Now lets import the html and extract correct tables: 

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

```

```{r}
nsduh_table_format(html = nsduh_html, table_num = 1)
nsduh_table_format(html = nsduh_html, table_num = 4)
nsduh_table_format(html = nsduh_html, table_num = 5)
```
```{r}
nsduh_df_1 = 
  tibble(
    drug = c("marj", "cocaine", "heroin"),
    table_n = c(1,4,5)
    ) |> 
  mutate(
    table = map(
      table_n, 
      nsduh_table_format, 
      html = nsduh_html)
  ) |> 
  unnest(table) 

nsduh_df_2 = 
  tibble(
    drug = c("marj", "cocaine", "heroin"),
    table_n = c(1,4,5)
    ) |> 
  mutate(
    table = map(
      table_n,
      \(x) nsduh_table_format(html = nsduh_html, table_num = x)))|> 
  unnest(table) 

```

Now we two ways to get a data frame with all of this information and we did so using map. Rather than having to use a for loop, we put a column into a data set and now we have a regular ole df. 

```{r}
new_york = 
  nsduh_df_2 |>  
  filter(State == "New York") 
```

### weather data example: 
```{r}

```

