---
title: "Iteration"
author: "Kaleb J. Frierson"
date: "2024-10-29"
output: html_document
---

# Notes

for loops:

Loops are the easiest place to start with iteration. Loops have a consistent output object. The basic structure: 

input = list(...)
output = list(...)

for (i in 1:n) {
output [[i]] = f(input[[i]])

}

Mapping across input list, running function, saving output. 

Loop functions make your life easier. You can use them in the context of a tidyverse like approach. Makes your code easier to read. More focus on the things you are trying to do/change. Map function is in the purrr package and is a loop function that helps clean up loops. Don't use base R loop functions, they're terrible per Jeff. 

output = map(input, f)

This produces the same result as a for loop but is better because: [answer here]

Map variants: 

By default, map makes one input and will return a list. If you know what kind of output your function will produce you can change default settings. 

Process: 

Write a single example for fixed inputs. 
Abstract example to a function
Embed in a loop 
Re-write useing a map statement 

(IE build the car model)

Lists: they exist, now we will worry about it. 

R uses lists to keep track of anything you want to use. They're a collection of things. One list can have lots of different things in them. 

Data frames - special kind of list. 

Lists can contain lists. If you start with a df and have a new list column and want to put it into a df, you can put it in there. This is very helpful. Inside of your data frame you can have input list as a column. Keeping everything in one data frame with list columns means there are fewer things to worry about. 

# Time to Code

```{r, message=FALSE}
library(tidyverse)
library(rvest)
```


## Here's some lists 

```{r}
l = list(
  vec_numeric = 1:4, 
  unif_sample = runif(100), 
  mat = matrix(1:8, nrow = 2, ncol = 4, byrow = TRUE), 
  summary = summary(rnorm(1000))
)

```

List more useful: 

```{r}
list_norm = 
  list(
    a = rnorm(20, 0, 5),
    b = rnorm(20, 4, 5), 
    c = rnorm(20, 0, 10), 
    d = rnorm(20, 4, 10)
  )

list_norm$b
```
lets reuse the function we wrote last time: 
```{r}
mean_and_sd = function(x) {
  
  mean_x = mean(x)
  sd_x = sd(x)
  
  out_df = 
    tibble(
      mean=mean_x, 
      sd = sd_x
    )
  
  return(out_df) 
}


```


Lets use the function to take the mean and sd of all samples: 

```{r}
mean_and_sd(list_norm$a)
mean_and_sd(list_norm$b)
mean_and_sd(list_norm$c)
mean_and_sd(list_norm$d) 
```



